---
title: "A C-R-A-F-T Walkthrough"
author: "Yuehong Cassandra Tai, Hyein Ko"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{A C-R-A-F-T Walkthrough}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment  = "#>",
  fig.width  = 7,
  fig.height = 4.5
)
library(craft)
```

> **Demo data, not real data.** The datasets used in this vignette
> are **synthetic but structurally based on the climate-stance
> demonstration used in the CRAFT paper** (Ko, Tai & Webb Williams).
> The CRAFT framework itself is general; climate stance is just one
> illustration. Class balance, error rates, and ideology
> distributions mimic the demonstration data, but no row corresponds
> to any actual legislator post. The five rationale-audit examples
> in the F step are lifted verbatim from the paper's Table
> `tab:example`. The real demonstration dataset will replace this
> demo when the paper is published.

This vignette walks through the five C-R-A-F-T steps. Labels are
binary (`support` / `opposing`); uncertainty is captured through
per-LLM confidence scores (R, A, F steps) and via the implicit
`1 - sup - opp` residual at the legislator level (T step), not as a
third label class.

## C — Construct role and task

```{r construct}
rt <- role(
  conception  = "annotator",
  task        = "classify climate stance of state legislators on X",
  gold        = TRUE,
  prompt_type = "zero-shot"
)
rt
```

The printed suggestions tell you which families of reliability and
validity metrics apply for this role-task combination.

## R — Report dual-track metrics

`reliab()` dispatches one of six inter-rater reliability metrics
(Cohen's kappa, weighted kappa, Fleiss' kappa, Krippendorff's alpha,
ICC, or percent agreement). `valid()` computes validity (precision,
recall, F1, MCC, etc.) against a gold standard. `dual()` bundles
both into one call.

```{r dual-track}
demo <- read.csv(
  system.file("extdata", "craft_demo.csv", package = "craft")
)
head(demo[, c("id", "gold_standard", "gpt5_label",
              "gemini3_label", "llama3_label")])

dt <- dual(
  ratings = demo[, c("gold_standard", "gpt5_label",
                     "gemini3_label", "llama3_label")],
  gold = demo$gold_standard,
  pred = list(                       # named list -> one row per
    "GPT-5"     = demo$gpt5_label,   # annotator in the validity table
    "Gemini-3"  = demo$gemini3_label,
    "Llama-3.3" = demo$llama3_label
  ),
  reliability_method = "kripp",      # >2 raters, nominal labels
  reliability_level  = "nominal",
  validity_metrics = c("f1_macro", "f1_weighted", "mcc",
                       "balanced_accuracy")
)
dt$reliability
dt$validity
```

`dual()` accepts either a single `pred` vector or a named list of
predictors. The list form (above) gives one row per annotator in the
validity table — useful when comparing several LLMs against the same
gold standard. The single-vector form is shown below:

```{r dual-single, eval=FALSE}
dual(ratings   = demo[, c("gold_standard", "gpt5_label")],
     gold      = demo$gold_standard,
     pred      = demo$gpt5_label,
     pred_name = "GPT-5",                # cleaner label in report()
     reliability_method = "cohen")
```

For pairwise Cohen's kappa across every annotator pair (the data
behind a kappa heatmap):

```{r pairwise}
reliab_pairs(
  demo[, c("gold_standard", "gpt5_label",
           "gemini3_label", "llama3_label")],
  method = "cohen",
  order  = c("gold_standard", "gpt5_label",
             "gemini3_label", "llama3_label")
)
```

## A — Assess stability

`stab()` computes reliability + validity for each model or prompt
variant separately, producing a tidy one-row-per-variant table:

```{r stability}
stab(
  predictions = list(
    "GPT-5"     = demo$gpt5_label,
    "Gemini-3"  = demo$gemini3_label,
    "Llama-3.3" = demo$llama3_label
  ),
  gold = demo$gold_standard,
  reliability_method = "cohen",
  validity_metrics   = c("f1_macro", "f1_weighted", "mcc",
                         "balanced_accuracy")
)
```

## F — Field audit and adjudication

`audit()` surfaces the rows that need human review by combining each
LLM's label with its confidence:

```{r audit}
# Pass text + rationales through so report() can quote them later
ann <- data.frame(
  id                = demo$id,
  text              = demo$text,
  label_a           = demo$gpt5_label,
  label_b           = demo$gemini3_label,
  confidence_a      = demo$confidence_gpt5,
  confidence_b      = demo$confidence_gemini3,
  rationale_gpt5    = demo$rationale_gpt5,
  rationale_gemini3 = demo$rationale_gemini3,
  stringsAsFactors  = FALSE
)

a <- audit(ann, confidence_threshold = 0.55, low_threshold = 0.31)
table(a$audit_status)
```

The five rows lifted from the paper's Table `tab:example` carry the
rationale text directly, so you can inspect why each annotator chose
a given label:

```{r rationale-audit}
paper_rows <- demo[demo$source == "main.tex Table tab:example",
                   c("id", "text", "llama3_label",
                     "rationale_gpt5", "rationale_llama3")]
paper_rows
```

Notice the Manchin/Reich row (`paper_ex_05`): GPT-5 and Gemini-3
labelled it `support` with a context-aware rationale; Llama-3.3
labelled it `opposing` after fixating on the surface phrases
`coal stocks` and `blocks efforts`. That is a real audit catch.

`tau_sens()` shows how the audit categorization shifts as the
confidence threshold moves:

```{r tau-sens}
tau_sens(ann, thresholds = c(0.5, 0.55, 0.6, 0.7))
```

## T — Translate to inference

The legislator-level dataset ships with the columns the DSL analysis
needs: `state`, outcomes (`sup`, `opp`), LLM predictions (`pred_sup`,
`pred_opp`), covariates, and the inclusion probability
(`cand_incl_prob_all`).

```{r load-legis}
load(system.file("extdata", "data.rda", package = "craft"))
str(data)
cat("Uncertainty residual: mean ",
    round(mean(1 - data$pred_sup - data$pred_opp), 3),
    ", median ",
    round(median(1 - data$pred_sup - data$pred_opp), 3),
    "\n", sep = "")
```

The implicit `1 - sup - opp` residual is small (mean ~4%), which
matches the climate-stance demonstration where most legislators have
clearly classifiable climate posts. **Practical consequence:** when
uncertainty is small, `dsl_cmp(sup ~ ...)` and `dsl_cmp(opp ~ ...)`
produce near-mirror-image coefficient estimates. When uncertainty is
more substantial, the two outcomes can have genuinely independent
driver patterns.

```{r translate-sup, eval=requireNamespace("dsl", quietly = TRUE)}
dsl_results <- dsl_cmp(
  data    = data,
  formula = sup ~ shor_ideo + per_mining + female + senate + perc_bchhigherE,
  predicted_var = "sup",
  prediction    = "pred_sup",
  sample_prob   = "cand_incl_prob_all",
  fixed_effect  = "oneway",
  index   = "state",
  cluster = "state"
)
dsl_results
```

The opposing-stance model is the symmetric specification:

```{r translate-opp, eval=requireNamespace("dsl", quietly = TRUE)}
dsl_cmp(
  data    = data,
  formula = opp ~ shor_ideo + per_mining + female + senate + perc_bchhigherE,
  predicted_var = "opp",
  prediction    = "pred_opp",
  sample_prob   = "cand_incl_prob_all",
  fixed_effect  = "oneway",
  index   = "state",
  cluster = "state"
)
```

## Emit a reproducibility report

`report()` writes a prose, methods-supplement-style document
capturing the LLM, version, prompt, threshold decisions, metrics,
audit examples, and DSL results. Output goes to whatever path you
pass; relative paths resolve against your current working directory.
The shipped `prompt_dev4.txt` is the actual dev4 codebook used in
the climate-stance demonstration of the CRAFT paper.

> **Long prompts.** In HTML output the prompt is rendered inside a
> collapsible `<details>` block — readers click to expand. In PDF
> output (where there is no toggle) the prompt is rendered as a
> wrapped verbatim block so the codebook is fully visible for
> transparency.

### Default: HTML in your working directory

```{r report-html, eval=FALSE}
report(
  role_task   = rt,
  llm         = "GPT-5",
  version     = "gpt-5-2025-09-15",
  prompt_path = system.file("extdata", "prompt_dev4.txt", package = "craft"),
  metrics     = dt,
  audit       = a,
  dsl         = dsl_results,
  thresholds  = list(confidence = 0.55, low = 0.31),
  output      = "craft_report.html"      # writes to getwd()
)
```

### PDF for paper supplement (requires LaTeX)

```{r report-pdf, eval=FALSE}
report(
  role_task   = rt,
  llm         = "GPT-5",
  version     = "gpt-5-2025-09-15",
  prompt_path = system.file("extdata", "prompt_dev4.txt", package = "craft"),
  metrics     = dt,
  audit       = a,
  dsl         = dsl_results,
  thresholds  = list(confidence = 0.55, low = 0.31),
  output      = "craft_report.pdf"
)
# First time only:
# tinytex::install_tinytex()
```

### Both formats side-by-side

```{r report-both, eval=FALSE}
report(
  role_task   = rt,
  llm         = "GPT-5",
  version     = "gpt-5-2025-09-15",
  prompt_path = system.file("extdata", "prompt_dev4.txt", package = "craft"),
  metrics     = dt,
  audit       = a,
  dsl         = dsl_results,
  thresholds  = list(confidence = 0.55, low = 0.31),
  output      = "craft_report",          # no extension
  format      = "both"                   # writes .html AND .pdf
)
```

### Customizing the DSL table for non-climate studies

The report's DSL table groups predictors as **Individual level** and
**District level** by default, with labels tuned to the
climate-stance demonstration. For other studies, attach
`term_labels` and `term_groups` attributes to the data frame
returned by `dsl_cmp()` before passing it to `report()`:

```{r dsl-customize, eval=FALSE}
my_dsl <- dsl_cmp(
  data    = my_data,
  formula = outcome ~ treatment + age + income + urban,
  predicted_var = "outcome",
  prediction    = "pred_outcome",
  sample_prob   = "incl_prob",
  fixed_effect  = "oneway",
  index   = "region",
  cluster = "region"
)

attr(my_dsl, "term_labels") <- c(
  treatment = "Treatment",
  age       = "Age",
  income    = "Household income",
  urban     = "Urban residence"
)
attr(my_dsl, "term_groups") <- list(
  "Primary"  = c("treatment"),
  "Controls" = c("age", "income", "urban")
)

report(..., dsl = my_dsl)
```

The rendered report contains prose narration of all CRAFT steps,
the prompt (toggle in HTML, expanded in PDF), a table of dual-track
metrics with one validity row per annotator, the stability table,
1-2 illustrative audit cases drawn from rows in your `audit` data
frame that carry `text` and `rationale_*` columns, and the
grouped Original-vs-DSL coefficient table.

## Where to go next

- `vignette("getting-started")` — five-minute introduction.
- `?reliab` — list of reliability methods and when to use each.
- `?report` — full argument reference for the reproducibility
  report generator.
