---
title: "Getting Started with craft (5-minute tutorial)"
author: "Yuehong Cassandra Tai, Hyein Ko"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Getting Started with craft (5-minute tutorial)}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>")
library(craft)
```

> **Demo data, not real data.** The `craft_demo.csv` file used below
> is **synthetic but structurally based on the climate-stance
> demonstration used in the CRAFT paper** (Ko, Tai & Webb Williams).
> Climate stance is used to illustrate the framework — CRAFT itself is
> domain-general. Five of the rows carry text and rationales lifted
> verbatim from the paper's Table `tab:example`; the rest are
> synthetic label sequences with realistic error rates. No row
> corresponds to any unpublished real annotation.

This tutorial gets you from zero to a complete CRAFT evaluation in
about five minutes. For the deeper end-to-end example (including DSL
correction), see `vignette("craft-walkthrough")`.

## What you need

A data frame with one column per rater (annotator or LLM), a
gold-standard column, and optionally per-rater confidence scores.

```{r demo-data}
demo <- read.csv(
  system.file("extdata", "craft_demo.csv", package = "craft")
)
head(demo)
```

Columns: `id`, `gold_standard`, `gpt5_label`, `gemini3_label`,
`llama3_label`, plus a `confidence_*` column per LLM. Labels are
binary (`support` / `opposing`); uncertainty is captured through
per-LLM confidence scores (R, A, F steps) and via the implicit
`1 - sup - opp` residual at the legislator level (T step), not as a
third label class.

## Step 1 (C): Document the role

```{r role}
rt <- role(conception  = "annotator",
           task        = "classify climate stance",
           gold        = TRUE,
           prompt_type = "zero-shot")
rt
```

## Step 2 (R): Compute reliability + validity

Pick the reliability metric that fits your data. For four categorical
raters (one human + three LLMs), Krippendorff's alpha is appropriate:

```{r reliab}
reliab(demo[, c("gold_standard", "gpt5_label",
                "gemini3_label", "llama3_label")],
       method = "kripp", level = "nominal")
```

Pairwise Cohen's kappa - every rater pair separately:

```{r reliab-pairs}
reliab_pairs(demo[, c("gold_standard", "gpt5_label",
                      "gemini3_label", "llama3_label")],
             method = "cohen")
```

Validity vs. gold (you pick the metrics you care about):

```{r valid}
valid(gold = demo$gold_standard,
      pred = demo$gpt5_label,
      metrics = c("f1_macro", "f1_weighted", "mcc", "accuracy"))
```

Both in one call:

```{r dual}
# Single annotator vs. gold
dual(
  ratings   = demo[, c("gold_standard", "gpt5_label")],
  gold      = demo$gold_standard,
  pred      = demo$gpt5_label,
  pred_name = "GPT-5",
  reliability_method = "cohen",
  validity_metrics   = c("f1_macro", "f1_weighted", "mcc")
)
```

To get validity for **all three annotators in one call**, pass a
named list of prediction vectors:

```{r dual-multi}
dual(
  ratings = demo[, c("gold_standard", "gpt5_label",
                     "gemini3_label", "llama3_label")],
  gold = demo$gold_standard,
  pred = list(
    "GPT-5"     = demo$gpt5_label,
    "Gemini-3"  = demo$gemini3_label,
    "Llama-3.3" = demo$llama3_label
  ),
  reliability_method = "kripp",
  validity_metrics   = c("f1_macro", "f1_weighted", "mcc")
)
```

## Step 3 (A): Compare across models

```{r stab}
stab(
  predictions = list(
    "GPT-5"     = demo$gpt5_label,
    "Gemini-3"  = demo$gemini3_label,
    "Llama-3.3" = demo$llama3_label
  ),
  gold = demo$gold_standard
)
```

## Step 4 (F): Audit

```{r audit}
ann <- data.frame(
  id           = demo$id,
  label_a      = demo$gpt5_label,
  label_b      = demo$gemini3_label,
  confidence_a = demo$confidence_gpt5,
  confidence_b = demo$confidence_gemini3
)
a <- audit(ann, confidence_threshold = 0.55, low_threshold = 0.31)
table(a$audit_status)
```

## Step 5 (T): DSL correction

`dsl_fit()` and `dsl_cmp()` propagate misclassification uncertainty
into a downstream regression. See `vignette("craft-walkthrough")` for
the full T-step example.

## Emit a reproducibility report

The report is written in prose and can render to HTML or PDF. The
output file lands in your current working directory by default, or
at any path you specify. The PDF output requires LaTeX (one-time
`tinytex::install_tinytex()`).

```{r report, eval=FALSE}
report(
  role_task   = rt,
  llm         = "GPT-5",
  version     = "gpt-5-2025-09-15",
  prompt_path = system.file("extdata", "prompt_dev4.txt", package = "craft"),
  metrics     = dual(
    ratings = demo[, c("gold_standard", "gpt5_label")],
    gold    = demo$gold_standard,
    pred    = demo$gpt5_label,
    reliability_method = "cohen"
  ),
  thresholds  = list(confidence = 0.55, low = 0.31),
  output      = "craft_report.html"      # writes to getwd()
)

# Or generate both an HTML preview AND a PDF for a paper supplement:
report(..., output = "craft_report", format = "both")
```

## Where to go next

- `vignette("craft-walkthrough")` — the deeper example with DSL.
- `?reliab` — full list of reliability methods and when to use each.
- `?report` — argument reference for the reproducibility report.
