---
title: "Producing a Transparency Report for Journal Submission"
author: "Yuehong Cassandra Tai, Hyein Ko"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Producing a Transparency Report for Journal Submission}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment  = "#>",
  fig.width  = 7,
  fig.height = 4.5
)
library(craft)
```

## Why this matters

Replication and transparency norms for studies that use large
language models are still emerging. Different journals will adopt
different disclosure requirements over time, and the most defensible
choice authors can make is to document the LLM pipeline thoroughly
even when no specific requirement is in place yet. Reviewers and
future readers benefit from knowing the exact model and version, the
verbatim prompts, the metrics that were computed against
gold-standard human labels, the threshold or decision rules applied,
the audit decisions made, and any correction applied for
measurement error in downstream analyses.

`craft::report()` bundles all of these into a single
methods-supplement-style document that can be circulated with
coauthors, archived alongside replication materials, or uploaded
with a journal submission when called for. The same call produces
both an **HTML** version (with the prompt behind a toggle, easy to
share by email) and a **PDF** version (clean and printable for
archival use).

> **What this vignette covers.** A minimal end-to-end recipe: from
> existing LLM outputs and a gold-standard audit subset to a
> rendered transparency document. For more conceptual depth on each
> CRAFT step see `vignette("craft-walkthrough")`.

## Minimum inputs you need

To generate a publishable report you need:

1. **A prompt text file** — the exact instruction issued to the model,
   saved as `.txt`. Don't paraphrase it. Whatever you actually sent
   the API, that is what goes in this file.
2. **LLM outputs** — model labels and (ideally) per-prediction
   confidence scores for every annotated case.
3. **A gold-standard audited subset** — at least a few hundred
   human-coded cases against which to compute validity.
4. **A DSL result** (if your downstream analysis is regression-based)
   — generated with `dsl_cmp()`.

Everything else flows from these.

## A walked example using the package demo

The package ships synthetic demo data plus the actual codebook used
in the climate-stance demonstration of the CRAFT paper.

### Step 1 — describe the role of the LLM

```{r role}
rt <- role(
  conception  = "annotator",
  task        = "classify the climate stance of state legislators",
  gold        = TRUE,
  prompt_type = "zero-shot"
)
rt
```

The printed summary is exactly what will appear at the top of the
generated report.

### Step 2 — compute the metrics that go into the report

```{r metrics}
demo <- read.csv(
  system.file("extdata", "craft_demo.csv", package = "craft")
)

dt <- dual(
  ratings = demo[, c("gold_standard", "gpt5_label",
                     "gemini3_label", "llama3_label")],
  gold = demo$gold_standard,
  pred = list(
    "GPT-5"     = demo$gpt5_label,
    "Gemini-3"  = demo$gemini3_label,
    "Llama-3.3" = demo$llama3_label
  ),
  reliability_method = "kripp",
  validity_metrics = c("f1_macro", "f1_weighted", "mcc",
                       "balanced_accuracy")
)
```

The validity table will appear with one row per annotator. For a
single-annotator study just pass a vector instead of a list — see
`?dual` for the alternative form.

### Step 3 — record the audit decisions

```{r audit}
ann <- data.frame(
  id                = demo$id,
  text              = demo$text,
  label_a           = demo$gpt5_label,
  label_b           = demo$gemini3_label,
  confidence_a      = demo$confidence_gpt5,
  confidence_b      = demo$confidence_gemini3,
  rationale_gpt5    = demo$rationale_gpt5,
  rationale_gemini3 = demo$rationale_gemini3,
  stringsAsFactors  = FALSE
)
a <- audit(ann, confidence_threshold = 0.55, low_threshold = 0.31)
```

The report quotes 1-2 illustrative cases from this object. Passing
`text` and `rationale_*` columns is what enables qualitative
examples of how human auditors used the model rationales to
adjudicate borderline labels — useful for any reader trying to
understand how the annotation pipeline made its decisions.

### Step 4 — run the DSL correction

```{r dsl, eval=requireNamespace("dsl", quietly = TRUE)}
load(system.file("extdata", "data.rda", package = "craft"))
dsl_results <- dsl_cmp(
  data    = data,
  formula = sup ~ shor_ideo + per_mining + female + senate + perc_bchhigherE,
  predicted_var = "sup",
  prediction    = "pred_sup",
  sample_prob   = "cand_incl_prob_all",
  fixed_effect  = "oneway",
  index   = "state",
  cluster = "state"
)
```

### Step 5 — emit the transparency document

```{r report, eval=FALSE}
report(
  role_task   = rt,
  llm         = "GPT-5",
  version     = "gpt-5-2025-09-15",
  prompt_path = system.file("extdata", "prompt_dev4.txt", package = "craft"),
  metrics     = dt,
  audit       = a,
  dsl         = dsl_results,
  thresholds  = list(confidence = 0.55, low = 0.31),
  output      = "craft_transparency_report",   # no extension
  format      = "both"                          # writes .html AND .pdf
)
```

The HTML file is for sharing with coauthors during preparation. The
PDF is suitable for archival use or for inclusion in supplementary
materials when relevant.

## What the rendered report contains

Each numbered section corresponds to one CRAFT step:

| Section | Content |
|---|---|
| **1. Construct role-task (C)** | How the LLM was conceptualized and what task it performed |
| **2. Model and prompt identification** | Model name, version string, and the verbatim prompt |
| **3. Threshold decisions** | Confidence cutoffs or other decision rules applied to outputs |
| **4. Report dual-track metrics (R)** | Reliability + validity tables (one row per annotator when multiple LLMs are compared) |
| **5. Assess stability (A)** | Cross-prompt or cross-model variability of the reported metrics |
| **6. Field audit (F)** | Distribution of audit statuses plus 1-2 illustrative cases with text and rationales |
| **7. Translate to inference (T)** | Original vs. DSL-corrected coefficient table |
| **8. Reproducibility note** | Timestamp + citation block |

If a section's inputs are not supplied (for example, no DSL was
run), the section prints a brief note instead of an empty table.

## A pre-submission checklist for thoroughness

- [ ] The model version string is the **exact API identifier**, not
  a marketing name. ("GPT-5" → `gpt-5-2025-09-15`.)
- [ ] The prompt file is **verbatim**, including system message,
  any few-shot examples, and JSON-schema constraints.
- [ ] The threshold values shown in Section 3 match the decision
  rules described in your Methods section.
- [ ] The audit examples in Section 6 represent the kinds of cases
  reviewers will want to inspect (boundary cases, model
  disagreements, common-error patterns).
- [ ] The DSL columns in Section 7 use the same outcome and
  covariates as in your main regression table.

## Citing `craft` in your methods section

A short example of language you can adapt:

> "We used GPT-5 (`gpt-5-2025-09-15`) and Gemini-3-Flash to
> independently annotate posts for climate stance. Inter-rater
> reliability between human and model annotators was assessed with
> Krippendorff's alpha, and validity with macro-F1, weighted F1, and
> MCC. Disagreements between models were adjudicated by a human
> coder following the protocol described in our transparency
> supplement, generated using the `craft` R package (Tai & Ko,
> 2026). Misclassification uncertainty in the AI-generated labels
> was propagated into the regression estimates via design-based
> supervised learning (Egami et al., 2023; implemented in
> `craft::dsl_cmp()`)."

## Where to go next

- `?report` — full argument reference for the report function.
- `vignette("craft-walkthrough")` — the deeper conceptual tour.
- `vignette("getting-started")` — five-minute introduction.
