---
title: "CRAFT Reproducibility Report"
date: "`r format(Sys.time(), '%Y-%m-%d')`"
params:
  role_task: NULL
  llm: NULL
  version: NULL
  prompt: NULL
  metrics: NULL
  audit: NULL
  dsl: NULL
  thresholds: NULL
  timestamp: NULL
output:
  html_document:
    toc: true
    toc_depth: 2
  pdf_document:
    toc: true
    toc_depth: 2
    latex_engine: xelatex
header-includes:
  - \usepackage{threeparttable}  # for the notes block under the DSL table
  - \usepackage{makecell}        # for stacked estimate / (SE) cells
  - \usepackage{booktabs}        # for \toprule / \midrule / \bottomrule
  - \usepackage{pifont}          # \ding{51} = checkmark (font-safe)
  - \newcommand{\cmark}{\ding{51}}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE,
                      results = "asis")

rt <- params$role_task

# Article-style words for the conception
conception_phrase <- switch(
  rt$conception,
  annotator = "like-human annotators",
  ml_system = "machine learning systems",
  silicon   = "silicon participants",
  rt$conception
)

# Gold-label phrase
gold_phrase <- if (isTRUE(rt$gold)) {
  "with a human-coded gold-standard subset for evaluation"
} else {
  "without a human-coded gold-standard subset"
}

prompt_phrase <- switch(
  rt$prompt_type,
  `zero-shot`  = "zero-shot prompting",
  `few-shot`   = "few-shot prompting",
  `fine-tuned` = "fine-tuned prompting",
  rt$prompt_type
)
```

*This report was generated by the `craft` R package implementing the
C-R-A-F-T framework (Ko, Tai & Webb Williams). It is intended as a
methods-transparency supplement to accompany the parent paper.*

# 1. Construct the role-task (C)

In this study, large language models are used as
**`r conception_phrase`** for the task of **`r rt$task`**,
`r gold_phrase`, using **`r prompt_phrase`**.

`r rt$suggested_metrics$notes`

# 2. Model and prompt identification

The annotations were produced by **`r params$llm`**, version
`r sprintf('\x60%s\x60', params$version)`. The exact prompt used is
reproduced verbatim below for replicability.

```{r prompt-block, results = "asis"}
# Render the prompt inside a code block. For PDF this is a plain
# \verbatim environment (no fancyvrb/listings needed, works on any
# LaTeX install). For HTML it becomes a fenced <pre> block.
#
# Two preprocessing steps for the PDF path:
#   1. Transliterate Unicode chars whose glyphs may be missing from
#      the verbatim font (≤ → <=, em-dash → --, etc.).
#   2. Wrap long lines to ~78 chars so they don't overflow the page
#      margin. Line structure within the prompt is preserved; only
#      overlong lines get wrapped.
.transliterate <- function(x) {
  x <- gsub("≤", "<=",  x)
  x <- gsub("≥", ">=",  x)
  x <- gsub("→", "->",  x)
  x <- gsub("←", "<-",  x)
  x <- gsub("—", "--",  x)
  x <- gsub("–", "-",   x)
  x <- gsub("…", "...", x)
  x <- gsub("[‘’]", "'", x)
  x <- gsub("[“”]", '"', x)
  x
}
.wrap_long_lines <- function(x, width = 78) {
  lines <- strsplit(x, "\n", fixed = TRUE)[[1]]
  wrapped <- unlist(lapply(lines, function(ln) {
    if (nchar(ln) <= width) ln else strwrap(ln, width = width)
  }), use.names = FALSE)
  paste(wrapped, collapse = "\n")
}

if (knitr::is_latex_output()) {
  prepped <- .wrap_long_lines(.transliterate(params$prompt), width = 78)
  cat("\\begin{verbatim}\n", prepped, "\n\\end{verbatim}\n", sep = "")
} else {
  # HTML: collapse the prompt behind a toggle. The <details> element is
  # closed by default; readers click to expand and inspect the verbatim
  # text. Helps when the prompt is hundreds of lines.
  n_lines <- length(strsplit(params$prompt, "\n", fixed = TRUE)[[1]])
  cat('<details>\n',
      '<summary><strong>Show full prompt</strong>',
      sprintf(' (%d lines)</summary>\n\n', n_lines),
      '```\n', params$prompt, '\n```\n',
      '\n</details>\n', sep = "")
}
```

# 3. Threshold decisions

```{r thresholds}
if (!is.null(params$thresholds)) {
  th <- params$thresholds
  parts <- vapply(names(th), function(k) {
    sprintf("**%s** = %s", k, format(th[[k]], digits = 3))
  }, character(1))
  cat("The following decision parameters were applied:\n\n",
      paste("-", parts, collapse = "\n"), "\n", sep = "")
} else {
  cat("No threshold decisions were recorded for this run.\n")
}
```

# 4. Report dual-track metrics (R)

```{r metrics}
m <- params$metrics
if (is.null(m)) {
  cat("_No metrics supplied._\n")
} else if (is.list(m) && all(c("reliability", "validity") %in% names(m))) {
  cat("Inter-rater reliability across all rater columns was assessed ",
      "with **", m$reliability$method[1], "**, yielding a value of **",
      sprintf("%.3f", m$reliability$value[1]), "**.\n\n", sep = "")

  if ("annotator" %in% names(m$validity)) {
    # Multi-annotator validity table
    ann_list <- m$validity$annotator
    cat("Validity metrics compare each annotator (",
        paste(sprintf("**%s**", ann_list), collapse = ", "),
        ") against the human gold-standard labels.\n\n", sep = "")
    cat(knitr::kable(m$validity, digits = 3,
                     caption = "Validity metrics by annotator vs. gold standard"),
        sep = "\n")
  } else {
    pred_label <- if (!is.null(m$pred_name)) m$pred_name else "the model output"
    cat("Validity metrics compare **", pred_label,
        "** against the human gold-standard labels.\n\n", sep = "")
    cat(knitr::kable(m$validity, digits = 3,
                     caption = sprintf(
                       "Validity metrics: %s vs. gold standard",
                       pred_label)),
        sep = "\n")
  }
  cat("\n\n")
} else {
  cat("Cross-model stability is reported below.\n\n")
  cat(knitr::kable(m, digits = 3,
                   caption = "Stability across model / prompt variants"),
      sep = "\n")
  cat("\n\n")
}
```

# 5. Assess stability (A)

```{r stability}
m <- params$metrics
is_stab <- is.data.frame(m) && "variant" %in% names(m)
if (is_stab) {
  span <- if ("f1_macro" %in% names(m)) range(m$f1_macro, na.rm = TRUE) else NULL
  cat("Stability was assessed across ", nrow(m),
      " model or prompt variants. ",
      if (!is.null(span)) sprintf("F1-macro ranges from %.3f to %.3f.",
                                  span[1], span[2]) else "",
      "\n\n", sep = "")
  cat(knitr::kable(m, digits = 3), sep = "\n")
  cat("\n\n")
} else {
  cat("Cross-prompt or cross-model stability comparisons were not ",
      "supplied with this report. The CRAFT framework recommends that ",
      "at least two prompt variants and at least two model variants be ",
      "evaluated on the same validation subset.\n", sep = "")
}
```

# 6. Field audit (F)

```{r audit}
a <- params$audit
if (is.null(a)) {
  cat("_No audit results supplied._\n")
} else {
  summary_tbl <- as.data.frame(table(audit_status = a$audit_status),
                               stringsAsFactors = FALSE)
  names(summary_tbl) <- c("audit_status", "n")
  cat("A total of **", nrow(a), "** annotations were audited. ",
      "The distribution of audit outcomes is reported below.\n\n", sep = "")
  cat(knitr::kable(summary_tbl,
                   caption = "Audit status distribution"),
      sep = "\n")
  cat("\n\n")

  # Show 1-2 example rows with text/rationale if those columns exist
  text_cols <- intersect(c("text", "label_a", "label_b",
                           "rationale_gpt5", "rationale_gemini3",
                           "rationale_llama3"), names(a))
  has_text <- "text" %in% names(a) && any(nchar(a$text) > 0, na.rm = TRUE)
  if (has_text) {
    ex <- a[nchar(a$text) > 0, , drop = FALSE]
    ex <- head(ex, 2)
    cat("\n\n**Illustrative cases.** The two cases below show how human ",
        "auditors used the model rationales to adjudicate borderline ",
        "labels.\n\n", sep = "")
    for (i in seq_len(nrow(ex))) {
      cat("> **Case ", i, ".** ", ex$text[i], "\n>\n", sep = "")
      if (!is.null(ex$label_a)) {
        cat("> - Annotator A label: ", ex$label_a[i], "\n", sep = "")
      }
      if (!is.null(ex$label_b)) {
        cat("> - Annotator B label: ", ex$label_b[i], "\n", sep = "")
      }
      if ("rationale_gpt5" %in% names(ex) && nchar(ex$rationale_gpt5[i]) > 0) {
        cat("> - GPT-5 rationale: ", ex$rationale_gpt5[i], "\n", sep = "")
      }
      if ("rationale_gemini3" %in% names(ex) && nchar(ex$rationale_gemini3[i]) > 0) {
        cat("> - Gemini-3 rationale: ", ex$rationale_gemini3[i], "\n", sep = "")
      }
      cat("\n")
    }
  }
}
```

# 7. Translate to inference (T)

```{r dsl}
d <- params$dsl
if (is.null(d)) {
  cat("_No design-based (DSL) correction was supplied._\n")
} else {

  # ---------------------------------------------------------------
  # Pretty term labels + row grouping.
  #
  # Default mapping covers the climate-stance demonstration. Users
  # can override by attaching attributes to the data frame passed
  # via `dsl`:
  #   attr(dsl, "term_labels") <- c(my_var = "My Variable", ...)
  #   attr(dsl, "term_groups") <- list(
  #     "Individual level" = c("var1", "var2"),
  #     "District level"   = c("var3", "var4")
  #   )
  # ---------------------------------------------------------------
  default_labels <- c(
    shor_ideo       = "Ideology",
    female          = "Female",
    senate          = "Senate",
    per_mining      = "Fossil fuel industry",
    perc_bchhigherE = "Education"
  )
  default_groups <- list(
    "Individual level" = c("shor_ideo", "female", "senate"),
    "District level"   = c("per_mining", "perc_bchhigherE")
  )

  term_labels <- attr(d, "term_labels")
  if (is.null(term_labels)) term_labels <- default_labels

  term_groups <- attr(d, "term_groups")
  if (is.null(term_groups)) term_groups <- default_groups

  # Drop intercepts and any term we don't have a label for
  d <- d[!grepl("Intercept", d$term), ]
  d$label <- ifelse(d$term %in% names(term_labels),
                    term_labels[d$term], d$term)

  # Significance stars
  d$stars <- ifelse(d$p.value < 0.01,  "***",
             ifelse(d$p.value < 0.05, "**",
             ifelse(d$p.value < 0.10, "*", "")))

  # Reshape long -> wide: rows = term, cols = "Original" / "DSL"
  models <- unique(d$model)

  # cell_for: returns "<est><stars>SEP(<se>)" where SEP is "\\\\" for LaTeX
  # (which becomes "\\" after R string parsing -> the LaTeX line break)
  # and "<br>" for HTML markdown.
  cell_for <- function(term, model, sep) {
    row <- d[d$term == term & d$model == model, , drop = FALSE]
    if (nrow(row) == 0) return("--")
    sprintf("%.3f%s%s(%.3f)", row$estimate[1], row$stars[1],
            sep, row$std.error[1])
  }

  # Use only terms we have labels for, in the order term_groups dictates
  ordered_terms <- unlist(term_groups, use.names = FALSE)
  ordered_terms <- ordered_terms[ordered_terms %in% d$term]

  cat("The table below reports the regression coefficients before ",
      "(Original) and after (DSL) applying the design-based correction ",
      "for measurement error in the LLM-generated labels.\n\n", sep = "")

  if (knitr::is_latex_output()) {
    # ----- LaTeX: grouped, threeparttable, checkmark FE rows -----
    cat("\\begin{table}[htbp!]\n",
        "\\centering\n",
        "\\caption{Original vs.\\ DSL-corrected estimates}\n",
        "\\begin{threeparttable}\n",
        "\\renewcommand{\\arraystretch}{1.15}\n",
        "\\setlength{\\tabcolsep}{8pt}\n",
        "\\begin{tabular}{l", paste(rep("c", length(models)), collapse = ""), "}\n",
        "\\toprule\n",
        "\\textbf{Predictor} & ", paste(models, collapse = " & "), " \\\\\n",
        "\\midrule\n", sep = "")
    for (grp in names(term_groups)) {
      grp_terms <- intersect(term_groups[[grp]], ordered_terms)
      if (length(grp_terms) == 0) next
      cat("\\multicolumn{", 1 + length(models), "}{l}{\\textit{",
          grp, "}} \\\\\n", sep = "")
      for (tm in grp_terms) {
        cells <- vapply(models,
                        function(m) sprintf("\\makecell{%s}",
                                            cell_for(tm, m, sep = "\\\\")),
                        character(1))
        cat("\\quad ", term_labels[tm], " & ",
            paste(cells, collapse = " & "), " \\\\\n", sep = "")
      }
      cat("\\addlinespace[4pt]\n")
    }
    cat("\\midrule\n",
        "State fixed effects & ",
        paste(rep("\\cmark", length(models)), collapse = " & "),
        " \\\\\n",
        "State-clustered SEs & ",
        paste(rep("\\cmark", length(models)), collapse = " & "),
        " \\\\\n",
        "\\bottomrule\n",
        "\\end{tabular}\n",
        "\\begin{tablenotes}[para,flushleft]\n",
        "\\small\n",
        "\\item \\textit{Notes}: Cells report coefficient estimates with ",
        "standard errors in parentheses. $^{*}p<0.10$, $^{**}p<0.05$, ",
        "$^{***}p<0.01$.\n",
        "\\end{tablenotes}\n",
        "\\end{threeparttable}\n",
        "\\end{table}\n\n", sep = "")
  } else {
    # ----- HTML: grouped markdown table -----
    header <- c("| Predictor | ",
                paste(models, collapse = " | "), " |")
    sep    <- c("|:--|", paste(rep(":---:", length(models)), collapse = "|"), "|")
    cat(header, "\n", sep = "")
    cat(sep, "\n", sep = "")
    for (grp in names(term_groups)) {
      grp_terms <- intersect(term_groups[[grp]], ordered_terms)
      if (length(grp_terms) == 0) next
      cat("| _", grp, "_ | ",
          paste(rep("", length(models)), collapse = " | "),
          " |\n", sep = "")
      for (tm in grp_terms) {
        cells <- vapply(models,
                        function(m) cell_for(tm, m, sep = "<br>"),
                        character(1))
        cat("| &nbsp;&nbsp;", term_labels[tm], " | ",
            paste(cells, collapse = " | "), " |\n", sep = "")
      }
    }
    cat("| **State fixed effects** | ",
        paste(rep("&check;", length(models)), collapse = " | "), " |\n", sep = "")
    cat("| **State-clustered SEs** | ",
        paste(rep("&check;", length(models)), collapse = " | "), " |\n\n", sep = "")
    cat("<sub>*Notes:* Cells report coefficient estimates with standard ",
        "errors in parentheses. *p<0.10, **p<0.05, ***p<0.01.</sub>\n\n",
        sep = "")
  }
}
```

# 8. Reproducibility note

This report was generated on **`r params$timestamp`** by the `craft`
R package. Re-running the same pipeline with the same data, prompt,
and threshold decisions should reproduce these numbers exactly.

When using the `craft` package or the CRAFT framework, please cite:

> Ko, H., Tai, Y. C. & Webb Williams, N. (2026). Can We Trust
> LLM-Generated Data? The CRAFT Framework for Measurement and
> Inference in Political Science.

> Tai, Y. C. & Ko, H. (2026). craft: A CRAFT Pipeline for Evaluating
> LLM-Generated Data. R package version 0.0.0.9000.
> https://github.com/casstai/craft-r
